---
title: "Final Project Wine Quality"
author: "Lara Vartanian 804633245"
date: "12/14/2017"
output: html_document
---

```{r}
wine_complete <- read.csv("~/Documents/Stats 101A /Final Project/Wine 2017 Training Complete.csv")
attach(wine_complete)
```

# Splitting data into 70/30
```{r}
# 70% of the sample size
sample_size <- floor(0.70 * nrow(wine_complete))

# Set seed to make the partition reproduceable
set.seed(12345) # better to set odd seed
train_ind <- sample(seq_len(nrow(wine_complete)), size = sample_size)

#train data has 70% of the original dataset 
train <- wine_complete[train_ind, ]
#test data has 30% of the original dataset 
test <- wine_complete[-train_ind, ]

# Qual has the original Quality values
Qual <- test$Quality  # so this is the y 

# Predicted has predicted Quality of the 10% of the dataset 
predicted_yvalues <- predict(previous_working_2, test) # this is my y-hat 

# Check the SSE for the 10% of training data 
(errors_for_70_30 <- sum((Qual - predicted_yvalues)^2))
```

Attempt 1: doing MLR with all predictors with no transformations yet 
```{r}
m1 <- lm(Quality ~ alcohol + density + pH + volatile.acidity + total.sulfur.dioxide + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + sulphates + as.numeric(Wine.Color))
summary(m1)
anova(m1) # top 4 SSreg contributors: density, pH, volatile.acidity, total.sulfur.dioxide
par(mfrow = c(2,2))
plot(m1)

library(corrplot)
cormat <- round(cor(wine_complete[,unlist(lapply(wine_complete, is.numeric))], use = "pairwise.complete.obs"), 3)
cormat[1,]
library(car)
scatterplotMatrix(cormat)
par(mfrow=c(1,1))
corrplot.mixed(cormat, lower = "number", upper = "pie")

pairs(Quality ~ alcohol + density + pH + volatile.acidity + total.sulfur.dioxide + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + sulphates)

vif(m1)
#removing density 
summary(powerTransform(cbind(Quality, alcohol, pH, volatile.acidity, total.sulfur.dioxide, residual.sugar, chlorides, free.sulfur.dioxide, sulphates)~1))# took out citric acid & density 

```

# T R A N S F O R M A T I O N 
```{r}
m1_minus_denisty <- lm(Quality ~ alcohol + pH + volatile.acidity + total.sulfur.dioxide + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + sulphates + as.numeric(Wine.Color))
summary(m1_minus_denisty)
# alcohol, pH, volatile aciditiy, total sulfur dioxide, residual sugar, free sulfur dioxide and sulphates all are super significant 

library(MASS)
par(mfrow=c(1,1))
boxcox(m1_minus_denisty)

summary(powerTransform(cbind(Quality, alcohol, pH, volatile.acidity, total.sulfur.dioxide, residual.sugar, chlorides, free.sulfur.dioxide, sulphates)~1))# took out citric acid & density 


m2 <- lm(Quality ~ I(alcohol^(-2)) + I(pH^(-1.5)) + log(volatile.acidity) + sqrt(total.sulfur.dioxide) + log(residual.sugar) + I(chlorides^(-.5)) + I(free.sulfur.dioxide^(.25)) + I(sulphates^(-.5)), data = train)
summary(m2)
# SSE based on m2 is 3183.162

m2_1 <- lm(Quality ~ I(alcohol^(-2)) + I(pH^(-1.5)) + log(volatile.acidity) + I(total.sulfur.dioxide^(.75)) + log(residual.sugar) + I(chlorides^(-.25)) + I(free.sulfur.dioxide^(.25)) + I(sulphates^(-.5)), data = train)
summary(m2_1)
vif(m2_1)
# SSE = 3189

mstd <- lm(Quality ~ alcohol + pH + volatile.acidity + total.sulfur.dioxide + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + sulphates)
par(mfrow = c(2,2))
plot(m2)
summary(m2)
vif(m2)
library(alr3)
par(mfrow=c(2,2))
mmp(m2,I(alcohol^(-2)),key=NULL)
mmp(m2,I(pH^(-1.5)),key=NULL)
mmp(m2,log(volatile.acidity),key=NULL)
mmp(m2,sqrt(total.sulfur.dioxide),key=NULL)
mmp(m2,log(residual.sugar),key=NULL)
mmp(m2,citric.acid,key=NULL)
mmp(m2,I(chlorides^(-.5)),key=NULL)
mmp(m2,I(free.sulfur.dioxide^(.25)),key=NULL)
mmp(m2,I(sulphates^(-.5)),key=NULL)
mmp(m2,mstd$fitted.values,xlab="Fitted Values",key=NULL)
# fits better
```
alcohol, pH, volatile aciditiy, total sulfur dioxide, residual sugar, free sulfur dioxide and sulphates all are super significant 


# P O L Y N O M I A L  R E G R E S S I O N 
I want to see if a polynomial regression gives a better model, so I attempt to create a polynomial regression, applying poly to the significant variables to see if my R-squared increases relative to the model with transformations suggested with function powerTransform
```{r}
model_try_2 <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T)+ poly(total.sulfur.dioxide, 3, raw = T) + poly(pH, 3, raw = T) + poly(residual.sugar, 1, raw = T) + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T), data = train)
summary(model_try_2)
vif(model_try_2)
# total sulfur dioxide and free sulfue dioxide have VIF > 5. This leads to multicollinearity problem
# SSE = 3126.072

m2_poly <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T)+ poly(pH, 3, raw = T) + poly(residual.sugar, 3, raw = T) + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T), data = train)
summary(m2_poly)  
# SSE = 3292.134

model_try_2.7 <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T) + poly(pH, 3, raw = T) + residual.sugar + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T), data = train)
summary(model_try_2.7)
# SSE = 3142
vif(model_try_2.7)  # VIF < 5 

# SSE = 3135 if sulphates is 1 

# TRY with diff see
# 3176 with poly 3 
# 3177  with poly 1 

# 3190 wuth suplates 2 
# 3182 with sulphates 1 
# 


# add another variable or not ? Should i add chlorides ? use anova

m2_poly_1 <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T) + poly(pH, 3, raw = T) + residual.sugar + poly(free.sulfur.dioxide, 3, raw = T) + poly(chlorides, 3, raw = T) + poly(sulphates, 3, raw = T), data = train)
summary(m2_poly_1)
# SSE = 3144 
vif(m2_poly_1)  # VIF < 5
anova(m2_poly_1)

# Check if we should add chlorides or not 
anova(model_try_2.7, m2_poly_1) 
# it is not significant so I will drop variable chlorides 
```

Adding Interactions
```{r}
model_interact <- lm(Quality ~ .*., data = train)
summary(model_interact)
# significant interactions: Wine.ColorW:citric.acid, Wine.ColorW:density, Wine.ColorW:alcohol, citric.acid:free.sulfur.dioxide
# Super significant interactions: Wine.ColorW:pH, fixed.acidity:chlorides, volatile.acidity:pH, chlorides:pH, chlorides:alcohol

model_interact_1 <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T)+ poly(pH, 3, raw = T) + residual.sugar + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T), alcohol:Wine.Color, data = train)
summary(model_interact_1)


model_try_2.7_4 <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T)+ poly(pH, 3, raw = T) + residual.sugar + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T) + alcohol:Wine.Color + Wine.Color:pH, data = train)  
summary(model_try_2.7_4)
vif(model_try_2.7_4)


model_interact_2 <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T)+ poly(pH, 3, raw = T) + residual.sugar + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T), poly(alcohol, 3, raw = T):Wine.Color, data = train)
summary(model_interact_2)
# SSE = 3142
vif(model_try_2.7)  # VIF < 5 



model_interact_m2 <- lm(Quality ~ I(alcohol^(-2)) + I(pH^(-1.5)) + log(volatile.acidity) + sqrt(total.sulfur.dioxide) + log(residual.sugar) + I(chlorides^(-.5)) + I(free.sulfur.dioxide^(.25)) + I(sulphates^(-.5)) + Wine.Color:I(pH^(-1.5)) + I(alcohol^(-2)):Wine.Color, data = train)
summary(model_interact_m2)
# SSE = 3150.155
vif(model_interact_m2)
par(mfrow=c(2,2))
plot(model_interact_m2)

#3164: no transformed interaction
# SSE: 3192


previous_working <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T)+ poly(pH, 3, raw = T) + residual.sugar + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T) + poly(alcohol, 3, raw = T):Wine.Color + Wine.Color:poly(pH, 3, raw = T), data = train)
summary(previous_working)
vif(previous_working)  # GVIF^(1/(2*Df)) < 5 except for interaction term between poly pH and wine color which is expected since it is an interaction 

# SSE: 3127.134

previous_working_1 <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T)+ poly(pH, 3, raw = T) + residual.sugar + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T) + poly(alcohol, 3, raw = T):Wine.Color + Wine.Color:poly(pH, 3, raw = T) + poly(volatile.acidity, 3, raw = T): poly(pH, 3, raw = T), data = train)
summary(previous_working_1)
vif(previous_working_1) # look at GVIF^(1/(2*Df)) to compare across dimensions

# SSE 3131 

anova(previous_working, previous_working_1)
# interaction between poly volatile acidity and poly pH should not be dropped 

par(mfrow=c(2,2))
plot(previous_working_1)

previous_working_2 <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T)+ poly(pH, 3, raw = T) + residual.sugar + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T) + poly(alcohol, 3, raw = T):Wine.Color + Wine.Color:poly(pH, 3, raw = T) + poly(volatile.acidity, 3, raw = T): poly(pH, 3, raw = T) + chlorides:poly(pH, 3, raw = T), data = train)
summary(previous_working_2)
# SSE = 3130 

anova(previous_working_1, previous_working_2)
# can drop the interaction between chlorides and pH 


# ALSO WANTED TO CHECK SSE OF INTERACTIONS WITH THE TRANSOFRMATION SUGGESTED WITH POWERTRANSFORM 

model_interact_m3 <- lm(Quality ~ I(alcohol^(-2)) + I(pH^(-1.5)) + log(volatile.acidity) + sqrt(total.sulfur.dioxide) + log(residual.sugar) + I(chlorides^(-.5)) + I(free.sulfur.dioxide^(.25)) + I(sulphates^(-.5)) + Wine.Color:I(pH^(-1.5)) + I(alcohol^(-2)):Wine.Color + log(volatile.acidity):I(pH^(-1.5)), data = train)
summary(model_interact_m3)

# 3193


model_interact_m4 <- lm(Quality ~ I(alcohol^(-2)) + I(pH^(-1.5)) + log(volatile.acidity) + sqrt(total.sulfur.dioxide) + log(residual.sugar) + I(chlorides^(-.5)) + I(free.sulfur.dioxide^(.25)) + I(sulphates^(-.5)) + Wine.Color:I(pH^(-1.5)) + alcohol:Wine.Color + volatile.acidity:pH + chlorides:pH, data = train)
summary(model_interact_m4)
vif(model_interact_m4)
# 3149 SSE 

```

# W O R K I N G  &  F I N A L  M O D E L 
```{r}
previous_working_1 <- lm(Quality ~ poly(alcohol, 3, raw = T)+ poly(volatile.acidity, 3, raw = T)+ poly(pH, 3, raw = T) + residual.sugar + poly(free.sulfur.dioxide, 3, raw = T) + poly(sulphates, 3, raw = T) + poly(alcohol, 3, raw = T):Wine.Color + Wine.Color:poly(pH, 3, raw = T) + poly(volatile.acidity, 3, raw = T): poly(pH, 3, raw = T), data = train)
summary(previous_working_1)
vif(previous_working_1) # look at GVIF^(1/(2*Df)) to compare across dimensions

# SSE 3131 

par(mfrow=c(2,2))
plot(previous_working_1)

```


check correlation plot to see if normal distribution  















